---
title: "Prediction algorithm for given data"
author: "Alokik"
date: "16/11/2020"
output: html_document
---

# Downloading and saving of data
```{r echo=TRUE, message=FALSE}
# download the training data 
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile = "training.csv")

# download the test data
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile = "testing.csv")


# loading required libraries
library(caret)
library(e1071)


# uploading the training and testing data
set.seed(1234)
Traindata <- read.csv("training.csv", na.strings=c("NA","#DIV/0!", ""))
Testdata <- read.csv("testing.csv", na.strings=c("NA","#DIV/0!", ""))
```
# Exploaratory analysis and preprocessing of data
```{r echo=TRUE}
# Delete columns with all missing values
Traindata<-Traindata[,colSums(is.na(Traindata)) == 0]
Testdata <-Testdata[,colSums(is.na(Testdata)) == 0]

# Delete variables that are irrelevant to our current project: user_name, raw_timestamp_part_1, raw_timestamp_part_,2 cvtd_timestamp, new_window, and  num_window (columns 1 to 7). 
Traindata   <-Traindata[,-c(1:7)]
Testdata <-Testdata[,-c(1:7)]
```


# Cross validation
For Cross validation the given training data will be divided into subsets of 2.
  1. training subset of 70% of original training data.
  2. testing subset of 30% of original training data.
  
```{r echo=TRUE}

inTrain <- createDataPartition(Traindata$classe, p = 0.75, list = FALSE)

# subset
training <- Traindata[inTrain, ]
training$classe <- as.factor(training$classe)
testing <- Traindata[-inTrain, ]
testing$classe <- as.factor(testing$classe)
```

## Model building

There will be two models, one is regression tree and another is Random forest
```{r echo=TRUE}

# regression tree
library(rpart)
modDT <- train(classe~., data = training, method = "rpart")
predDT <-  predict(modDT, testing)

DT <- confusionMatrix(testing$classe, predDT)


# linear discrimination analysis
modRF <- train(classe~., data = training, method = "rf")
predRF <- predict(modRF, testing)

RF <- confusionMatrix(testing$classe, predRF)

data.frame(regression_tree= DT$overall["Accuracy"], Random_forest = RF$overall["Accuracy"])
```
**From the above result we can say randam forest method performed better than regression tree with accuracy 0.9945 (95% CI : (0.992, 0.9964)) compared to Regression Tree model with 0.4945 (95% CI : (0.4804, 0.5086)). The Random forest method is chosen for analysis. The expected out-of-sample error is estimated at 0.0065.**


# Submission
The result of Test data is:
```{r}
predict(modRF, Testdata)

```




